The first high-redshift Changing Look Quasars

### 2019-12-04 tex/CIV_CLQs.tex
quite a bit update on the plot/hexbin_plots/MBH_REW/ figures and code

\author[Ross {\it et al.}]
{Nicholas~P.~Ross$^{1}$\thanks{E-mail: npross@roe.ac.uk},
 Matthew Graham$^{2}$, K. E. Saavik Ford$^{3,4,5}$,  Barry McKernan$^{3,4,5}$   
\newauthor Daniel Stern$^{6}$ and Giorgio Calderone$^{7}$
\\
%%  List of institutions
$^{1}$Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill, Edinburgh EH9 3HJ, United Kingdom \\
$^{2}$Cahill Center for Astronomy and Astrophysics, California Institute of Technology, Mail Code 249/17, 1200 E California Blvd, Pasadena CA 91125, USA\\
$^{3}$Department of Science, BMCC, City University of New York, New York, NY 10007, USA \\
$^{4}$Department of Astrophysics, Rose Center for Earth and Space, American Museum of Natural History, Central Park West at 79th Street, NY 10024, USA \\
$^{5}$Graduate Center, City University of New York, 365 5th Avenue, New York, NY 10016, USA\\
$^{6}$Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Mail Stop 169-221, Pasadena, CA 91109, USA \\
$^{7}$INAF -- Osservatorio Astronomico di Trieste, Via Tiepolo 11, I-34143 Trieste, Italy \\
}

@@

For J1205+3422, our spectral observations cover 5195 days observed,
1691 days in the rest-frame. This quasar was initially identified in
SDSS in 2005 May, as a bright, $g\approx18.0$, blue-sloped quasar with
broad \siiv, \civ, \ciii and \mgii. \ciii and \civ\ have large
blueshifts of $\approx$2600$\pm$150 and $\approx$1150$\pm$100 km
s$^{-1}$, respectively.  By 2019, however, the optical brightness
dropped by $\sim$1.5 magnitudes and the spectra are significantly less
blue.  While \lya and \nv are detectable in both 2019 spectra, \civ\
has all but disappeared in the 2019 June spectrum.  The broad \ciii
emission has disappeared between the 2005 and 2019 spectra. The
changes in \civ\ and \ciii going from broad emission to barely
detectable have on the timescales of $\approx$50 days in the
rest-frame.

For J1638+2827, our spectral observations cover 4030 days observed,
1265 days in the rest-frame. Here, in the initial epoch spectrum,
\civ\ is broad and bright, as is C\,{\sc iii}. However, $\approx$400
rest-frame days later, the broad \civ\ and \ciii BEL have faded, the
continuum slope around 1400\AA\ has changed from $\approx-1.48$ to
$\approx-2.25$, but the \lya/\nv emission complex is very similar in
shape and line flux intensity. Around 870 days in the rest-frame after
the second spectral epoch, \lya, \nv, \civ, \ciii and \mgii are all
apparent and broad, with \mgii being seen for the first time at high
signal-to-noise. The light curve is consistent with this spectral
brightening, increasing from $\sim$20th magnitude to $\sim$18.5
magnitude at optical wavelengths. An absorption feature between \lya
and \nv is seen in all three spectral epochs.

@@

%% trim={<left> <lower> <right> <upper>}
\begin{figure}
  \centering
  \includegraphics[width=8.40cm, trim=0.37cm 0.3cm 0.0cm 0.2cm, clip]{figures/CIVregions.png}
    \vspace{-18pt}
  \caption{Observed spectra and best fit model in the region relevant
    to \civ\ emission line for the quasars considered here.
    The solid blue line is the Lorentzian profile fit the to the \civ\ emission line.
    The long-dashed red line is the continuum fit, with the solid orange
    line giving the overall fit. The short-dashed vertical lines gives the
    rest wavelengths of the \civ\ doublet. }
  \label{fig:QSFit-CIV}
\end{figure}
\subsection{\civ\ Emission Line Evolution}
We analyzed the spectra of the three quasars, at all observational
epochs, using the QSFit spectral fitting package
\citep{Calderone2017}.  One advantage of using QSFit is that it allows
constraints on the slope and luminosity of the broad band continuum of
the source. The relevant estimated quantities, including continuum
luminosity and slope at rest-frame 1450\AA, \civ\ line luminosity,
FWHM and EW are given in Tabl;e~\ref{tab:QSFit-results}. All fits are
performed with $E(B-V) = 0$ and the best fit model in the region of
the \civ\ emission line are shown in Figure~\ref{fig:QSFit-CIV}.

All \civ\ lines are fitted with a single, broad, Lorentzian profile.
This allows us to account for the narrow peak of the \civ\ line.  No
narrow components are considered for several reasons: in the epochs of
highest brightness the ``narrow'' component would have FWHM~$\sim2-3
\times 10^{3}$~km~s$^{-1}$, i.e. values exceeding the usual widths of
genuine narrow lines ($\lesssim$ $10^3$ km s$^{-1}$); by allowing a
second component, to have such large widths their parameters would
become highly degenerate with the ``broad'' components, and the latter
would also have much larger FWHM ($\sim$$10^{4}$ km s$^{-1}$); and by
neglecting the narrow components we have a consistent fit across all
epochs.

@@

\begin{figure}
  \centering
  \includegraphics[width=8.5cm, trim=0.2cm 0.2cm 0.0cm 0.2cm, clip]
  {figures/CIV_CLQs_MBHvsREW_QSFit20K.png}
   \vspace{-12pt}
   \caption[]{The virial black hole masses of $\approx$20,000 $z>1.5$ quasars
     from QSFit catalogue \citep{Calderone2017} and the \civ\ Equivalent widths.
     Different quasars are given different colours: J1205+3422 purple; J1638+2827 green; J2228+2201 is blue.
     Different epochs different symbols: first epoch circles; second epoch squares; third epoch diamonds. 
   }
  \label{fig:CIV_MBHvsREW}
\end{figure}
\subsection{Virial Black Hole Masses}\label{sec:BH_masses} 
The FWHM of broad lines is likely related to the mass of the
supermassive black hole powering the quasar phenomenon, which is
assumed to be constant on any human timescale.  Hence it is
instructive to check whether the virial product, which is the basic
quantity used to calculate the single epoch black hole mass estimate,
show any variation.

@@

\citet{Shen2011} reports Virial BH mass based on \mgii and \civ\ as
$\log (M/M_{\rm BH}) = 9.55\pm0.05$ and $\log(M/M_{\rm BH}) =
9.49\pm0.04$, respectively. We thus use the mean of these five values
(three from our three epochs and the two \citet{Shen2011} values) to
establish a working mean SMBH mass of $\log (M_{\rm BM, vir}/M_{\odot})
= 9.726$ for J1205+3422.  J1638+2827 also has two virial mass
measurments from Shen2011, but also has two measurements from
\citet{Kozlowski2017} (again using \civ and \mgii). The mean mass
measurement for J1638+2827 here is $\log(M_{\rm BM, vir}/M_{\odot} =
9.091$. For J2228+2201 there are two \citet{Kozlowski2017}, leading to
a mean mass of $\log(M_{\rm BM, vir}/M_{\odot} = 9.366$. We use these
mean SMBH masses when calculating the Eddington ratios.

@@

The velocity offsets of the \civ\ line are also
approximately constant, and compatible with a single value (within
3$\sigma$).  The exception is J2228+2201, where a significant change
($\sim$7$\sigma$) is observed between the second and third observation
epochs.

@@

\begin{table}
  \centering
  \begin{tabular}{l l l r}
    \hline
    \hline
    Object           & MJD      & $L_{\rm bol}$    &  $\eta_{\rm Edd}$  \\
    \hline
                         & 53498    & 47.216           &  -0.610                  \\
    J1205+3422 & 58538    &  46.567          &  -1.259  \\                        %% L_Edd (J12)  = 47.82637   for log(M) = 9.726
                         & 58693     &  46.070         &  -1.756 \\
    \hline 
                          & 54553   & 46.166    & -1.025 \\
    J1638+2827  & 55832   & 45.981    & -1.210  \\                                            %% L_Edd (J16)  = 47.191370  for log(M) = 9.091
                          & 58583   & 46.504   & -0.687       \\
    \hline 
                          & 56189   & 46.231   &  -1.235 \\
    J2228+2201   & 56960    & 47.342   & -0.124 \\                                             %% L_Edd (J22)  = 47.466370  or log(M) = 9.366
                          & 58693   & 46.826     &  -0.640 \\
    \hline
    \hline
  \end{tabular}
  \caption{
    Bolometric luminosities are given in either \citet{Shen2011} or 
    \citet{Kozlowski2017} for the three CLQs. We scale these 
    values using our measured $\nu L_{\nu}$ continuum luminosity values.
    From Section~\ref{sec:BH_masses}, we assume: $\log(M_{\rm BM, vir}/M_{\odot}) = 9.726$ for J1205+3422,
    $\log(M_{\rm BM, vir}/M_{\odot} = 9.091$ for  J1638+2827 
    and $\log(M_{\rm BM, vir}/M_{\odot} = 9.366$ for J2228+2201. 
    The Eddington Luminosity is 
    $L_{\rm Edd} = 1.26\times10^{38} (M/M_{\odot})$ erg s$^{-1}$ and
    %%  6.70456406676162      e+47  for J12
    %%  1.5537120898684228  e+47  for J16
    %%  2.9266483634099527  e+47 for J22
   $\eta_{\rm Edd}$ is the log of the Eddington ratio.}
\label{tab:Eddington_ratios} 
\end{table}

@@

\citet{Graham2019} report the largest number of H$\beta$ CSQs with 111
sources being identified. \citet{Graham2019} find that this population
of extreme varying quasars is associated with {\it changes} in the
Eddington ratio, (rather than just the magnitude of the Eddington
ratio itself) and the timescales imply cooling/heating fronts
propagating through the disk.\\
{\bf NPR to finish this off} \\

@@

\begin{figure*}
  \centering
  \includegraphics[width=14.5cm, trim=0.2cm 0.2cm 0.0cm 0.2cm, clip]
  {figures/MJD_vs_Eddington_20191202.png}
  \vspace{-12pt}
  \caption[]{Eddington Ratios of the three \civ\ CLQs.
    Grayscale gives the Eddington Ratio ranges for the SDSS
    \citep[from ][]{Shen2011} and the BOSS \citep[from ][]{Kozlowski2017)} 
    catalogues.
    Symbols as in Fig.~\ref{fig:CIV_MBHvsREW}.
    The red region ($L / L_{\rm Edd} \approx 0.02\pm0.01$;  $\eta =-1.7\pm0.13$) is
    the transition accretion rate suggested by \citet{NodaDone2018}.
    A ``High/Soft State'' with $\eta_{\rm Edd} \geq -1$ and a
    a ``Low/Hard State'' with $\eta_{\rm Edd} \leq -2$ are indicated 
    with dahsed lines and arrows. 
  }
  \label{fig:Eddington_ratios}
\end{figure*}
\subsection{Eddington Ratios and State Changes} 
The broad UV and optical lines in quasars are most sensitive to the
extreme ultraviolet (EUV) part of the spectral energy distribution
(SED), with \civ\ (and indeed \heii and \nv) being at the higher
energy end of the EUV distribution.

>>>>

\author[Ross {\it et al.}]
{Nicholas~P.~Ross$^{1}$\thanks{E-mail: npross@roe.ac.uk},
Matthew Graham$^{2}$,
Giorgio Calderone$^{3}$, 
K. E. Saavik Ford$^{4,5,6}$,  
\newauthor Barry McKernan$^{4,5,6}$  and Daniel Stern$^{7}$  
\\
%%  List of institutions
$^{1}$Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill, Edinburgh EH9 3HJ, United Kingdom \\
$^{2}$Cahill Center for Astronomy and Astrophysics, California Institute of Technology, Mail Code 249/17, 1200 E California Blvd, Pasadena CA 91125, USA\\
$^{3}$INAF -- Osservatorio Astronomico di Trieste, Via Tiepolo 11, I-34143 Trieste, Italy \\
$^{4}$Department of Science, BMCC, City University of New York, New York, NY 10007, USA \\
$^{5}$Department of Astrophysics, Rose Center for Earth and Space, American Museum of Natural History, Central Park West at 79th Street, NY 10024, USA \\
$^{6}$Graduate Center, City University of New York, 365 5th Avenue, New York, NY 10016, USA\\
$^{7}$Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Mail Stop 169-221, Pasadena, CA 91109, USA \\
}

@@

For J1205+3422, our spectral observations cover 5195 days observed,
1691 days in the rest-frame. This quasar was initially identified in
SDSS in 2005 May, as a bright, $g\approx18.0$, blue-sloped quasar with
broad Si\,{\sc iv}, \civ, \ciii and Mg\,{\sc ii}. \ciii and \civ\ have large
blueshifts of $\approx$2600$\pm$150 and $\approx$1150$\pm$100 km
s$^{-1}$, respectively.  By 2019, however, the optical brightness
dropped by $\sim$1.5 magnitudes and the spectra are significantly less
blue.  While \lya and \nv are detectable in both 2019 spectra, \civ\
has all but disappeared in the 2019 June spectrum.  The broad \ciii
emission has disappeared between the 2005 and 2019 spectra. The
changes in \civ\ and \ciii going from broad emission to barely
detectable have on the timescales of $\approx$50 days in the
rest-frame.

For J1638+2827, our spectral observations cover 4030 days observed,
1265 days in the rest-frame. Here, in the initial epoch spectrum,
\civ\ is broad and bright, as is C\,{\sc iii}. However, $\approx$400
rest-frame days later, the broad \civ\ and \ciii BEL have faded, the
continuum slope around 1400\AA\ has changed from $\approx-1.48$ to
$\approx-2.25$, but the \lya/\nv emission complex is very similar in
shape and line flux intensity. Around 870 days in the rest-frame after
the second spectral epoch, Ly$\alpha$, N\,{\sc v}, \civ, C\,{\sc iii}] and \mgii are all
apparent and broad, with \mgii being seen for the first time at high
signal-to-noise. The light curve is consistent with this spectral
brightening, increasing from $\sim$20th magnitude to $\sim$18.5th 
magnitude at optical wavelengths. An absorption feature between \lya
and \nv is seen in all three spectral epochs.

@@

%% trim={<left> <lower> <right> <upper>}
\begin{figure}
  \centering
  \includegraphics[width=8.40cm, trim=0.37cm 0.3cm 0.0cm 0.2cm, clip]{figures/CIVregions.png}
    \vspace{-18pt}
  \caption{Observed spectra and best fit model in the region relevant
    to \civ\ emission line for the quasars considered here.
    The solid blue line is the Lorentzian profile fit the to the \civ\ emission line.
    The long-dashed red line is the continuum fit, with the solid orange
    line giving the overall fit. The short-dashed vertical lines gives the
    rest wavelengths of the \civ\ doublet. }
  \label{fig:QSFit-CIV}
\end{figure}
\subsection{\civ\ Emission Line Evolution}
We analyzed the spectra of the three quasars, at all observational
epochs, using the QSFit spectral fitting package
\citep{Calderone2017}.  One advantage of using QSFit is that it allows
constraints on the slope and luminosity of the broad band continuum of
the source. The relevant estimated quantities, including continuum
luminosity and slope at rest-frame 1450\AA, \civ\ line luminosity,
FWHM and EW are given in Table~\ref{tab:QSFit-results}. All fits are
performed with $E(B-V) = 0$ and the best fit model in the region of
the \civ\ emission line are shown in Figure~\ref{fig:QSFit-CIV}.

All \civ\ lines are fitted with a single, broad, Lorentzian profile.
This allows us to account for the narrow peak of the \civ\ line.  No
narrow components are considered for several reasons: in the epochs of
highest brightness the ``narrow'' component would have FWHM~$\sim2-3
\times 10^{3}$~km~s$^{-1}$, i.e. values exceeding the usual widths of
genuine narrow lines ($\lesssim$10$^{3}$ km s$^{-1}$); by allowing a
second component, to have such large widths their parameters would
become highly degenerate with the ``broad'' components, and the latter
would also have much larger FWHM ($\sim$$10^{4}$ km s$^{-1}$); and by
neglecting the narrow components we have a consistent fit across all
epochs.

@@

\begin{figure}
  \centering
  \includegraphics[width=8.5cm, trim=0.2cm 0.2cm 0.0cm 0.2cm, clip]
  {figures/CIV_CLQs_MBHvsREW_QSFit20K.png}
   \vspace{-12pt}
   \caption[]{The virial black hole masses of $\approx$20,000 $z>1.5$ quasars
     from QSFit catalogue \citep{Calderone2017} and the \civ\
     Equivalent widths.  Different quasars are given different colours:
     J1205+3422 purple; J1638+2827 green; J2228+2201 is blue.  Different
     epochs different symbols: first epoch circles; second epoch squares;
     third epoch diamonds.  As noted in Table~\ref{tab:QSFit-results},
     $^{*}$ signifies the \civ\ line is very weak and the associated
     estimates are likely unreliable.}
   \label{fig:CIV_MBHvsREW}
\end{figure}
\subsection{Virial Black Hole Masses}\label{sec:BH_masses} 
The FWHM of broad lines is likely related to the mass of the
supermassive black hole powering the quasar phenomenon, which is
assumed to be constant on any human timescale.  Hence it is
instructive to check whether the virial product, which is the basic
quantity used to calculate the single epoch black hole mass estimate,
show any variation.

@@

\citet{Shen2011} reports Virial BH mass based on \mgii and \civ\ as
$\log (M_{\rm BH, Vir}/M_{\odot}) = 9.55\pm0.05$ and $\log(M_{\rm BH, Vir}/M_{\odot}) =
9.49\pm0.04$, respectively. We thus use the mean of these five values
(three from our three epochs and the two \citet{Shen2011} values) to
establish a working mean SMBH mass of $\log (M_{\rm BM, vir}/M_{\odot})
= 9.726$ for J1205+3422.  J1638+2827 also has two virial mass
measurments from Shen2011, but also has two measurements from
\citet{Kozlowski2017} (again using \civ and \mgii). The mean mass
measurement for J1638+2827 here is $\log(M_{\rm BM, vir}/M_{\odot} =
9.091$. For J2228+2201 there are two \citet{Kozlowski2017}, leading to
a mean mass of $\log(M_{\rm BM, vir}/M_{\odot} = 9.366$. We use these
mean SMBH masses when calculating the Eddington ratios.

@@

The velocity offsets of the \civ\ line are also
approximately constant, and compatible with a single value (within
3$\sigma$).  The exception is J2228+2201, where a significant change
($\sim$7$\sigma$) is observed between the second and third observation
epochs. The magnitude of the velocity blueshifts ($\sim200-1100$ km s$^{-1}$)
are consistent with recent results into the properties of 
rest-frame UV spectra over the redshift range $1.5 \leq z \leq 7.5$
\citep{Meyer2019}.

@@

\begin{table}
  \centering
  \begin{tabular}{l l l r}
    \hline
    \hline
    Object           & MJD      & $L_{\rm bol}$    &  $\eta_{\rm Edd}$  \\
    \hline
                     & 53498    & 47.216           &  -0.610                  \\
J1205+3422 & 58538    &  46.567          &  -1.259  \\                        %% L_Edd (J12)  = 47.82637   for log(M) = 9.726
                     & 58693     &  46.070         &  -1.756 \\
    \hline 
                     & 54553   & 46.166    & -1.025 \\
J1638+2827  & 55832   & 45.981    & -1.210  \\                                            %% L_Edd (J16)  = 47.191370  for log(M) = 9.091
                     & 58583   & 46.504   & -0.687       \\
    \hline 
                      & 56189   & 46.231   &  -1.235 \\
J2228+2201   & 56960    & 47.342   & -0.124 \\                                             %% L_Edd (J22)  = 47.466370  or log(M) = 9.366
                       & 58693   & 46.826     &  -0.640 \\
    \hline
    \hline
  \end{tabular}
  \caption{
    Bolometric luminosities are given in either \citet{Shen2011} or 
    \citet{Kozlowski2017} for the three CLQs. We scale these 
    values using our measured $\nu L_{\nu}$ continuum luminosity values.
    From Section~\ref{sec:BH_masses}, we assume: $\log(M_{\rm BM, vir}/M_{\odot}) = 9.726$ for J1205+3422; 
    $\log(M_{\rm BM, vir}/M_{\odot} = 9.091$ for  J1638+2827, 
    and $\log(M_{\rm BM, vir}/M_{\odot} = 9.366$ for J2228+2201. 
    The Eddington Luminosity is 
    $L_{\rm Edd} = 1.26\times10^{38} (M/M_{\odot})$ erg s$^{-1}$ and
    %%  6.70456406676162      e+47  for J12
    %%  1.5537120898684228  e+47  for J16
    %%  2.9266483634099527  e+47 for J22
   $\eta_{\rm Edd}$ is the log of the Eddington ratio.}
\label{tab:Eddington_ratios} 
\end{table}

@@

\citet{Graham2019b} report the largest number of H$\beta$ CSQs with 111
sources being identified. \citet{Graham2019b} find that this population
of extreme varying quasars is associated with {\it changes} in the
Eddington ratio, (rather than just the magnitude of the Eddington
ratio itself) and the timescales imply cooling/heating fronts
propagating through the disk.\\
{\bf NPR to finish this off} \\

@@

\begin{figure*}
  \centering
  \includegraphics[width=14.5cm, trim=0.2cm 0.2cm 0.0cm 0.2cm, clip]
  {figures/MJD_vs_Eddington_20191202.png}
  \vspace{-12pt}
  \caption[]{Eddington Ratios of the three \civ\ CLQs.
    Grayscale gives the Eddington Ratio ranges for the SDSS
    \citep[from ][]{Shen2011} and the BOSS \citep[from ][]{Kozlowski2017} 
    catalogues.
    Symbols as in Fig.~\ref{fig:CIV_MBHvsREW}.
    The red region ($L / L_{\rm Edd} \approx 0.02\pm0.01$;  $\eta =-1.7\pm0.13$) is
    the transition accretion rate suggested by \citet{NodaDone2018}.
    A ``High/Soft State'' with $\eta_{\rm Edd} \geq -1$ and a
    a ``Low/Hard State'' with $\eta_{\rm Edd} \leq -2$ are indicated 
    with dahsed lines and arrows. 
  }
  \label{fig:Eddington_ratios}
\end{figure*}
\subsection{Eddington Ratios and State Changes} 
The broad UV and optical lines in quasars are most sensitive to the
extreme ultraviolet (EUV) part of the spectral energy distribution
(SED), with \civ\ (and indeed \heii and \nv) being at the higher
energy end of the EUV distribution.
